---
title: "Untitled"
output: html_document
date: "2026-04-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# donut plot data for prism
* if a gene has multiple mechanisms, count for each mechanism once.
use all genes, without filtering
```{r}
library(tidyverse)

# read data
amr <- read.csv("all_AMR.csv", check.names = FALSE)

# define segment order (optional but recommended)
segment_order <- c(
  "Duodenum", "Jejunum", "Ileum",
  "AscendingColon", "TransverseColon", "DescendingColon"
)

# ---- split multiple mechanisms ----
amr_summary <- amr %>%
  filter(!is.na(Tissue), !is.na(`Resistance Mechanism`)) %>%
  separate_rows(`Resistance Mechanism`, sep = ";") %>%
  mutate(
    Tissue = factor(Tissue, levels = segment_order),
    `Resistance Mechanism` = str_trim(`Resistance Mechanism`)
  ) %>%
  count(Tissue, `Resistance Mechanism`, name = "Gene_count") %>%
  group_by(Tissue) %>%
  mutate(
    Total_mechanism_count = sum(Gene_count),
    Percent = Gene_count / Total_mechanism_count * 100
  ) %>%
  ungroup()

# ---- Prism input: counts (recommended) ----
prism_counts <- amr_summary %>%
  select(Tissue, `Resistance Mechanism`, Gene_count) %>%
  pivot_wider(
    names_from = `Resistance Mechanism`,
    values_from = Gene_count,
    values_fill = 0
  )

# ---- Prism input: percent (optional) ----
prism_percent <- amr_summary %>%
  select(Tissue, `Resistance Mechanism`, Percent) %>%
  pivot_wider(
    names_from = `Resistance Mechanism`,
    values_from = Percent,
    values_fill = 0
  )

# export
write.csv(prism_counts, "Prism_AMR_mechanism_counts_ALL.csv", row.names = FALSE)
write.csv(prism_percent, "Prism_AMR_mechanism_percent_ALL.csv", row.names = FALSE)
```
```

